### 2024-08-26 chapters/gaze_independence.tex
Add dropped gaze-independence literature

Gaze-independent ERP-based BCIs~\cite{Riccio2012} can be realized in three
ways.
Firstly, non-visual stimulation paradigms such as auditory and somatosensory paradigms
do not rely on gaze redirection but often result in lower information transfer
rates, increased mental effort and user-dependent variability~\cite{Reichert2020b}.
Secondly, visual stimulation can be optimized, e.g. such that the stimuli are always
present in the field of view~\cite{Treder2010,Pires2011,Lees2018}.
Non-spatial visual attention (feature attention) can also be exploited, such as
attention to stimulus color, shape or symbol~\cite{Zhang2010,Treder2011,Hwang2015}.
Alternative visual stimulation paradigms can modulate specific ERP components
that are more sensitive to stimulation in the visual periphery~\cite{Schaeff2012,Xu2022}.
%These methods, However, can suffer from reduced information
%transfer rates~\cite{Chennu2013}.
However, they still rely to some extent on
eye motor control, often necessitating central gaze fixation.

Firstly, by avoiding visual stimulation entirely and opting for auditory or
somatosensory stimulation instead.
However, these alternatives often result in lower information transfer
rates, increased mental effort and
user-dependent variability~\cite{Reichert2020b}.
Secondly, gaze-independence can be achieved through tailored stimulus presentation.
This can involve enforcing overt VSA by presenting stimuli that fill the field
of view of the user,
exploiting visual non-spatial attention, or a combination thereof~\cite{Treder2011}.
The most common example of non-spatial attention is feature attention, where
the user directs their attention to an expected property of some of the
presented stimuli, such as its color, its shape or the symbol it represents~\cite{Zhang2010,Hwang2015,Reichert2020a}.
In covert VSA scenarios, stimulus paradigms are often fine-tuned to modulate
specific ERP responses~\cite{Schaeff2012,Xu2022}, increasing the
signal-to-noise ratio (SNR).
Some noteworthy examples include the GIBS Block Speller~\cite{Pires2011},
the GeoSpell interface~\cite{Aloise2012a}, and the RSVP
speller~\cite{Acqualagna2011,Lees2018}.
These systems typically require users to focus on a central fixation point while
selecting peripheral targets.
However, they often suffer from reduced information
transfer rates~\cite{Chennu2013}.
Moreover, they still rely to some extent on
eye motor control, often necessitating central gaze fixation.

>>>>

Gaze-independent ERP-based BCIs~\cite{Riccio2012} can be realized in three
ways. Firstly, non-visual stimulation paradigms such as auditory and somatosensory paradigms
do not rely on gaze redirection but often result in lower information transfer
rates, increased mental effort and user-dependent variability~\cite{Reichert2020b}.

Secondly, visual stimulation can be optimized, e.g. such that the stimuli are always
present in the field of view either overtly~\cite{Acqualagna2013, Won2018,
Lin2018} or covertly~\cite{Treder2010,Pires2011,Lees2018}. Some noteworthy examples include the GIBS Block Speller~\parencite{Pires2011},
the GeoSpell interface~\parencite{Aloise2012}, and the RSVP
speller~\parencite{Acqualagna2011}.
Non-spatial visual attention (feature attention) can also be exploited, such as
attention to stimulus color, shape or symbol~\cite{Zhang2010,Treder2011,Hwang2015}.
Alternative visual stimulation paradigms can modulate specific ERP components
that are more sensitive to stimulation in the visual periphery~\cite{Schaeff2012,Xu2022}.
These systems typically require users to focus on a central fixation point while
selecting peripheral targets.
These methods, However, can suffer too from reduced information
transfer rates~\cite{Chennu2013}.
However, they still rely to some extent on
eye motor control, often necessitating central gaze fixation.
\todo{advantages
disadvantages rewrite (requires gaze fixation, low itr, not tested with split attention)}
